---
title: "ZiwanPeng_fraud"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


0.load library
```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(tidymodels)
library(janitor)
library(skimr)
library(glmnet)     # -- lasso and regularization 
library(rpart.plot) # -- plotting decision trees 
library(vip)   
library(reshape2)
library(lubridate)
```

1.read csv
```{r}
fraud <- read_csv("project_2_training.csv") %>%
  clean_names()
fraud
```

```{r}
fraud%>%
  skim()
```

```{r}
fraud <- fraud%>%
  mutate(event_label=as.factor(if_else(event_label=='fraud',1,0)))%>%
  mutate(year=year(event_timestamp),quarter=quarters(event_timestamp),month=month(event_timestamp),day=day(event_timestamp),weekday=weekdays(event_timestamp),hour=hour(event_timestamp))%>%
  mutate_if(is.character,as.factor)%>%
  mutate(across(.cols = c("month","year","day","quarter",'weekday','hour','billing_postal','event_id','card_bin'),as.factor))
  
```



```{r}
fraud %>%
  count(event_label)%>%
  mutate(pct=round(n/sum(n),4))


fraud %>%
  count(event_label)%>%
  mutate(pct = round(n/sum(n),4))%>%
  ggplot(aes(x=event_label,y=pct))+
  geom_col()+
  geom_text(aes(label=pct) ,color="red") + 
  labs(title = "Percentage of Fraud")
```

```{r}
fraud %>%
  skim()
```

###explore categorical variables

```{r}
fraud%>%
  group_by(event_label)%>%
  count(currency)%>%
  pivot_wider(values_from=c(n), names_from = event_label) %>%
  mutate(pct_1 = `1`/ (`0` + `1`),
             pct_0 = 1 - pct_1) 
```

1. find the percentage of fraud and non-fraud in email domain
2. filter the email domains that have pct_1 > 0.5
3. detect those suspicious email domains in all data set, define other email domains( whose pct_1 <0.5) as 'other'
```{r}
fraud%>%
  group_by(event_label)%>%
  count(email_domain)%>%
  pivot_wider(values_from=c(n), names_from = event_label) %>%
  mutate(pct_1 = `1`/ (`0` + `1`),
             pct_0 = 1 - pct_1) %>%
  filter(pct_1>0.5)%>%
  arrange(desc(pct_1))%>%
  mutate(email=email_domain)%>%
  select(email_domain,email)-> list_email


fraud<- left_join(fraud,list_email,by="email_domain")%>%
  mutate(email=as.character(email))%>%
  mutate(email=if_else(is.na(email),"other",email))%>%
  mutate(email=as.factor(email))

```
```{r}
fraud%>%
  group_by(event_label)%>%
  count(email)%>%
  ggplot(aes(x=email,y=n, fill=event_label))+
  geom_col(position="fill")+
  geom_hline(aes(yintercept=0.0543))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  labs(title="The suspicious email domains")
```

the suspicious billing postal code
```{r}
fraud%>%
  group_by(event_label)%>%
  count(billing_postal)%>%
  pivot_wider(values_from=c(n), names_from = event_label) %>%
  mutate(pct_1 = `1`/ (`0` + `1`),
             pct_0 = 1 - pct_1)%>%
  filter(pct_1>0.5)%>%
  mutate(billing_postal_code=billing_postal)%>%
  select(billing_postal_code,billing_postal)-> list_postal


fraud<- left_join(fraud,list_postal,by="billing_postal")%>%
  mutate(billing_postal_code=as.character(billing_postal_code))%>%
  mutate(billing_postal_code=if_else(is.na(billing_postal_code),"other",billing_postal_code))%>%
  mutate(billing_postal_code=as.factor(billing_postal_code))
```


```{r}
fraud%>%
  group_by(event_label)%>%
  count(billing_postal_code)%>%
  ggplot(aes(x=billing_postal_code,y=n, fill=event_label))+
  geom_col(position="fill")+
  geom_hline(aes(yintercept=0.0543))+
  labs(title="The suspicious billing postal code")
```
```{r}
fraud%>%
  group_by(event_label)%>%
  count(ip_address)%>%
  pivot_wider(values_from=c(n), names_from = event_label) %>%
  mutate(pct_1 = `1`/ (`0` + `1`),
             pct_0 = 1 - pct_1)%>%
  filter(pct_1>0.5)%>%
  mutate(ip=ip_address)%>%
  select(ip,ip_address)-> list_ip


fraud<- left_join(fraud,list_ip,by="ip_address")%>%
  mutate(ip=as.character(ip))%>%
  mutate(ip=if_else(is.na(ip),"other",ip))%>%
  mutate(ip=as.factor(ip))
```

```{r}
fraud%>%
  group_by(event_label)%>%
  count(ip)%>%
  ggplot(aes(x=ip,y=n, fill=event_label))+
  geom_col(position="fill")+
  geom_hline(aes(yintercept=0.0543))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  labs(title="The suspicious ip address")
```


```{r}

fac_v <- function(col){
  fraud%>%
  na.omit()%>%
  ggplot(aes(x=!!as.name(col),fill=event_label))+
  geom_bar(position = "fill")+
  geom_hline(aes(yintercept=0.0543))
}

fac_v('billing_state')
fac_v('currency')+labs(title="Fraud rate among currency")
fac_v("cvv")+labs(title="Fraud rate among CVV")
fac_v("signature_image")+labs(title="Fraud rate among signature image")
fac_v("transaction_type")
fac_v("transaction_env")+labs(title="Fraud rate in transaction environment")
fac_v("tranaction_initiate")+labs(title="Fraud rate in type of transaction initiation")
fac_v("year")
fac_v("quarter")
fac_v("month")
fac_v("day")
fac_v("weekday")+labs(title="Fraud rate in weekdays")
fac_v("hour")

```

###explore numeric variables
```{r}

num_v <- function(col){
  fraud%>%
  na.omit()%>%
  ggplot(aes(x=!!as.name(col),fill=event_label))+
  geom_histogram(bins=20,position = "fill")+
  geom_hline(aes(yintercept=0.0543))
}

for (col in colnames(fraud %>% select_if(is.numeric))){
 print(num_v(col))
 } 

num_v("transaction_adj_amt")+labs(title="Relationship between fraud rate and adjusted transaction amount")
num_v("inital_amount")+labs(title="Relationship between fraud rate and amount of first transaction")
num_v("account_age_days")+labs(title="Relationship between fraud rate and number of days since the account created")
```

###spliting
```{r}
set.seed(123)

train_test_spit<- initial_split(fraud, prop = 0.8, strata = event_label)

train <- training(train_test_spit)
test  <- testing(train_test_spit)

sprintf("Train PCT : %1.2f%%", nrow(train)/ nrow(fraud) * 100)
sprintf("Test  PCT : %1.2f%%", nrow(test)/ nrow(fraud) * 100)
```

###recipe
```{r}

fraud_recipe <- recipe(event_label ~ ., 
                      data = train) %>%
  step_rm(ip_address,user_agent,email_domain,phone_number,billing_city,billing_postal,
          card_bin,applicant_name,billing_address,merchant_id,locale,tranaction_initiate,
          year,quarter,month,day,weekday,hour,event_timestamp,days_since_last_logon,inital_amount)%>%
  update_role(event_id, new_role = "id variable")%>%
  step_novel(all_nominal_predictors()) %>%
  step_impute_median(all_numeric_predictors()) %>%
  step_impute_mode(all_nominal_predictors()) %>%
  step_normalize(all_numeric_predictors()) %>%
  step_dummy(all_nominal_predictors())
```
### logistic regression
```{r}
lg1<- logistic_reg(penalty = 0.001, mixture = 1) %>%
  set_mode("classification") %>%
  set_engine("glmnet")

logistic_wf <- workflow() %>%
  add_recipe(fraud_recipe) %>%
  add_model(lg1) %>%
  fit(train)

```


```{r}
logistic_wf %>%
  pull_workflow_fit() %>%
  tidy() %>%
   mutate(across(is.numeric,round,3))%>%
  arrange(desc(estimate))
```

```{r}
options(yardstick.event_first = FALSE)  
predict(logistic_wf, train, type="prob") %>%
  bind_cols(predict(logistic_wf, train, type="class")) %>%
  bind_cols(train)  %>%
  metrics(event_label, estimate = .pred_class, .pred_1)

predict(logistic_wf, test, type="prob") %>%
  bind_cols(predict(logistic_wf, test, type="class")) %>%
  bind_cols(test)  %>%
  metrics(event_label, estimate = .pred_class, .pred_1)

```
threshold
```{r}
predict(logistic_wf, test, type="prob")%>%
  bind_cols(test)  %>%
  pr_curve(event_label, .pred_1) %>%
  mutate(
    recall = round(recall, 2),
    .threshold = round(.threshold, 3),
    precision = round(precision, 3)
  ) %>%
  group_by(recall) %>%
  summarise(precision = max(precision),
            .threshold = min(.threshold))%>%
  mutate(false_positive_rate=round(6134*recall*(1/precision-1)/11849,2))%>%
  filter(false_positive_rate<=0.08 &false_positive_rate>=0.06)

```


```{r}
logistic_wf %>%
  pull_workflow_fit() %>%
  vip()+
  labs(title="importance of variables in lasso logistic regression model")
```


```{r}
predict(logistic_wf, train, type="prob") %>%
  mutate(.pred=as.factor(if_else(.pred_1>=0.395,1,0)))%>%
  bind_cols(train)  ->logistic_train
logistic_train%>%
  metrics(event_label, estimate = .pred, .pred_1)

predict(logistic_wf, test, type="prob") %>%
  mutate(.pred=as.factor(if_else(.pred_1>=0.395,1,0)))%>%
  bind_cols(test)   -> logistic_test
logistic_test%>%
  metrics(event_label, estimate = .pred, .pred_1)
```

```{r}
bind_rows(logistic_train %>%
  mutate(model = "train"), 
logistic_test %>%
  mutate(model = "test")) %>%
  group_by(model) %>%
  roc_curve(event_label, .pred_1) %>%
  autoplot() +
  geom_vline(xintercept=0.395, color="red") +
  labs(title = "ROC chart")+
  labs(title="ROC chart for lasso logistic regression model")
```

```{r}


logistic_train %>%
  conf_mat(event_label, estimate = .pred) %>%
  autoplot(type = "heatmap") + labs(title="confusion matrix in train")
logistic_test %>%
  conf_mat(event_label, estimate = .pred) %>%
  autoplot(type = "heatmap") + labs(title="confusion matrix in test")


```

```{r}
logistic_test %>%
  ggplot(aes(.pred_1, fill=event_label)) +
  geom_histogram(bins=100) +
  xlim(0, 1) +
  geom_vline(aes(xintercept=0.395)) +
  labs(title="change threshold to 0.395")
```





###ridge
```{r}
lg2<- logistic_reg(penalty = 0.001, mixture = 0) %>%
  set_mode("classification") %>%
  set_engine("glmnet")

ridge_wf <- workflow() %>%
  add_recipe(fraud_recipe) %>%
  add_model(lg2) %>%
  fit(train)
```

```{r}
options(yardstick.event_first = FALSE)  
predict(ridge_wf, train, type="prob") %>%
  bind_cols(predict(ridge_wf, train, type="class")) %>%
  bind_cols(train)  %>%
  metrics(event_label, estimate = .pred_class, .pred_1)

predict(ridge_wf, test, type="prob") %>%
  bind_cols(predict(ridge_wf, test, type="class")) %>%
  bind_cols(test)  %>%
  metrics(event_label, estimate = .pred_class, .pred_1)
```

tuning the threshold
```{r}
predict(ridge_wf, train, type="prob")%>%
  bind_cols(train)  %>%
  pr_curve(event_label, .pred_1) %>%
  mutate(
    recall = round(recall, 3),
    .threshold = round(.threshold, 3),
    precision = round(precision, 3)
  ) %>%
  group_by(recall) %>%
  summarise(precision = max(precision),
            .threshold = min(.threshold))%>%
  mutate(false_positive_rate=round(6134*recall*(1/precision-1)/11849,2))%>%
  filter(false_positive_rate==0.06)

```
##decision tree
```{r}
TP = 0
FP = 3
TN = 0
FN = 1
       
loss_matr <- matrix(c(TP, FP, FN, TN), nrow = 2, byrow = TRUE)
loss_matr

tree_spec <- decision_tree(cost_complexity = 0.0, tree_depth=30) %>%
  set_mode("classification") %>%
  set_engine("rpart", 
             parms = list(loss = loss_matr),
             model=TRUE)
```


```{r}
tree_wf <- workflow() %>%
  add_recipe(fraud_recipe) %>%
  add_model(tree_spec) %>%
  fit(train)
```


```{r}
options(yardstick.event_first = FALSE)
predict(tree_wf, train, type="prob") %>%
  bind_cols(predict(tree_wf, train, type="class")) %>%
  bind_cols(train)  %>%
  metrics(event_label, estimate = .pred_class, .pred_1)

predict(tree_wf, test, type="prob") %>%
  bind_cols(predict(tree_wf, test, type="class")) %>%
  bind_cols(test)  %>%
  metrics(event_label, estimate = .pred_class, .pred_1)
```

```{r}
predict(tree_wf, train, type="prob")%>%
  bind_cols(train)  %>%
  pr_curve(event_label, .pred_1) %>%
  mutate(
    recall = round(recall, 3),
    .threshold = round(.threshold, 3),
    precision = round(precision, 3)
  ) %>%
  group_by(recall) %>%
  summarise(precision = max(precision),
            .threshold = min(.threshold))%>%
  mutate(false_positive_rate=round(6134*recall*(1/precision-1)/11849,2))%>%
  filter(false_positive_rate==0.06)
```


##random forest
```{r}
forest_spec <-rand_forest(mtry=4, min_n=100, trees=100) %>%
  set_mode("classification")%>%
  set_engine("ranger", importance="impurity") 
```

```{r}
forest_wf <- workflow() %>%
  add_recipe(fraud_recipe) %>%
  add_model(forest_spec) %>%
  fit(train)
```

```{r}
options(yardstick.event_first = FALSE)
predict(forest_wf, train, type="prob") %>%
  bind_cols(predict(forest_wf, train, type="class")) %>%
  bind_cols(train)  %>%
  metrics(event_label, estimate = .pred_class, .pred_1)

predict(forest_wf, test, type="prob") %>%
  bind_cols(predict(forest_wf, test, type="class")) %>%
  bind_cols(test)  %>%
  metrics(event_label, estimate = .pred_class, .pred_1)
```


```{r}
predict(forest_wf, test, type="prob")%>%
  bind_cols(test)  %>%
  pr_curve(event_label, .pred_1) %>%
  mutate(
    recall = round(recall, 2),
    .threshold = round(.threshold, 3),
    precision = round(precision, 3)
  ) %>%
  group_by(recall) %>%
  summarise(precision = max(precision),
            .threshold = min(.threshold))%>%
  mutate(false_positive_rate=round(6134*recall*(1/precision-1)/11849,2))%>%
  filter(false_positive_rate==0.06)
```

###model comparison

```{r}
predict(logistic_wf, test, type="prob")%>%
  bind_cols(test)  -> log_test


predict(ridge_wf, test, type="prob")%>%
  bind_cols(test)  -> ridge_test

predict(tree_wf, test, type="prob")%>%
  bind_cols(test)  -> dt_test

predict(forest_wf, test, type="prob")%>%
  bind_cols(test)  -> rf_test



```




```{r}
bind_rows(log_test %>%mutate(model = "lasso"), ridge_test %>%mutate(model = "ridge"),dt_test %>%mutate(model = "decision tree"),rf_test %>%mutate(model = "random forest")) %>%
  group_by(model) %>%
  roc_curve(event_label, .pred_1) %>%
  autoplot() +
  labs(title="ROC chart for all models")
```


###kaggle
```{r}
fraud_kaggle <-read_csv("project_2_holdout.csv")%>%
  clean_names()%>%
  mutate(year=year(event_timestamp),quarter=quarters(event_timestamp),month=month(event_timestamp),day=day(event_timestamp),weekday=weekdays(event_timestamp),hour=hour(event_timestamp))%>%
  mutate_if(is.character,as.factor)%>%
  mutate(across(.cols = c("month","year","day","quarter",'weekday','hour','billing_postal','event_id','card_bin'),as.factor))

fraud_kaggle<- left_join(fraud_kaggle,list_email,by="email_domain")%>%
  mutate(email=as.character(email))%>%
  mutate(email=if_else(is.na(email),"other",email))%>%
  mutate(email=as.factor(email))


fraud_kaggle<- left_join(fraud_kaggle,list_postal,by="billing_postal")%>%
  mutate(billing_postal_code=as.character(billing_postal_code))%>%
  mutate(billing_postal_code=if_else(is.na(billing_postal_code),"other",billing_postal_code))%>%
  mutate(billing_postal_code=as.factor(billing_postal_code))


fraud_kaggle<- left_join(fraud_kaggle,list_ip,by="ip_address")%>%
  mutate(ip=as.character(ip))%>%
  mutate(ip=if_else(is.na(ip),"other",ip))%>%
  mutate(ip=as.factor(ip))


predict(forest_wf, fraud_kaggle, type="prob") %>%
  bind_cols(fraud_kaggle) -> kaggle

kaggle%>%
  mutate(event_id=as.character(event_id),event_label=.pred_1)%>%
  subset(select=c(event_id,event_label))%>%
  mutate(event_id=as.numeric(event_id))%>%
  write_csv("kaggle_lg.csv")
  

```



